---
title: "Gapminder analysis"
author: "Dan Marrone"
date: "10/4/2015"
output: html_document
---

This document will analyze some of the gapminder data. 

```{r, echo=FALSE}
#install.packages("dplyr")
library("dplyr")
```

```{r read-in-data, echo=FALSE}
gap.in<-read.table("output/combined_gapMinder.tsv", sep = "\t", header = TRUE)
#summary(gap.in)

gap.in %>% 
  select(pop,country,continent,year) %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarise(median = median(pop))
```

```{r plotting-fcn, echo=FALSE}

make_plot <- function(data,selctry) {
  subdata <- filter(data, country == selctry) 
  plot(x = subdata$year, 
       y = subdata$gdp/1e12, 
       main = paste0(selctry,": GDP evolution"), 
       xlab = "Year", 
       ylab = "GDP (T$)", 
       pch = 20, tck = 0.01)
#       cex = 5*(subdata[,3]/max(subdata[,3])))
  axis(4, tick = TRUE, labels = FALSE, tck = 0.01)
  axis(3, tick = TRUE, labels = FALSE, tck = 0.01)
#  dev.off()
}

gap.in %>% 
  select(country,year,pop,gdpPercap) %>%
  filter(country == "Canada"| country == "Mexico"| country == "Chile") %>%
  group_by(country) -> ctrydata
ctrydata<-mutate(ctrydata,gdp=pop*gdpPercap)

#ctrydataCA<-filter(ctrydata, country == "Canada") 

```

## Plots of GDP versus time

Notably, Mexico has a larger GDP than Canada, and Chile's GDP was badly hurt by the Pinochet years.

```{r, echo=FALSE}
make_plot(ctrydata,"Canada")  
make_plot(ctrydata,"Mexico")  
make_plot(ctrydata,"Chile")  
```

## Life Expectancies

```{r life-expectancies, echo=FALSE}
# min mean max for each continent

gap.in %>% 
  select(lifeExp,country,continent,year) %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarise(minPop = min(lifeExp), meanPop = mean(lifeExp), maxPop = max(lifeExp))

```

African life expectancies are the shortest in the world, for the simple analysis performed here. Africa has the shortest life expectancy of any continent checked, and has the shortest mean. 
